---
title: "Homework 02"
output: github_document
author: "Robert Han"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("UsingR")) install.packages("UsingR"); library(UsingR)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("dplyr")) install.packages("dplyr"); library(dplyr)
if (!require("e1071")) install.packages("e1071"); library(e1071)
if (!require("gplots")) install.packages("gplots"); library(gplots) 
if (!require("reshape2")) install.packages("reshape2"); library(reshape2) 
```


## The Pick et al. Dataset
The dataset that you will be exploring will be found in the Homework02 folder.
It was downloaded from the Dryad Data Repository at:
https://doi.org/10.5061/dryad.40jp4
The paper describing the analysis of this data can be found here:
https://doi.org/10.1086/688918

**Disentangling Genetic and Prenatal Maternal Effects on Offspring Size and Survival**
Joel L. Pick, Christina Ebneter, Pascale Hutter, and Barbara Tschirren
*The American Naturalist* 2016 188:6, 628-639

Data of body mass, size and survival throughout development of Japanese quail chicks originating from reciprocal crosses of divergent artificial selection lines for prenatal maternal investment


```{r Importing Data}
quail <- read_csv("pick_et_al.csv", trim_ws = TRUE)
head(quail)
```

The following description of the data was also provided:
Column Name - Description
"animal" - offspring ID XXXXXXXXXXXXX
"mother" - maternal ID XXXXXXXXXXXXXX
"father" - paternal ID XXXXXXXXXXXXXXX
"hatch.mass" - offspring hatchling mass (g)
"week2.mass" - offspring week 2 mass (g)
"week2.tarsus" - offspring week 2 tarsus length (mm)
"week4.mass" - offspring week 4 mass (g)
"week4.tarsus" - offspring week 4 tarsus length (mm)
"adult.tarsus" - offspring adult tarsus length (mm)
"egg.mass" - egg mass (g) that offspring originated from
"maternal.line" - selection line of the mother; H = high investment line, L = low investment line
"paternal.line" - selection line of the father; H = high investment line, L = low investment line
"sex" - offspring sex
"replicate" - selection line replicate
"hatching.day" - day of hatching (17 or 18) relative to start of incubation (day 0)
"hatch.env" - hatching environment (specific incubator)
"week2.env" - chick rearing environment (specific cage) from hatching to week 2
"week4.env" - juvenile rearing environment (specific cage) from week 2 to week 4
"survive" - survival to adulthood
"exclude.surv" - whether individual was excluded from survival analysis or not (TRUE = excluded); see methods

```{r Summary}
summary(quail)
```

You can see that some of the data is categorical, but it is of the class "character". 
We should change that to ensure that it these variables are handled properly during analysis.
I have shown you how to do one of these, but you will need to complete this process, referring to the description of each variable.

```{r Factoring data}
library(dplyr)
quail$sex <- as.factor(quail$sex)
table(quail$sex)
```

Next explore the data to determine if variables look like they will need to be transformed or if they already have a normal distribution.
Note the importance of testing for normality for each population separately. 

```{r Data Exploration}
simple.eda(quail$hatch.mass)
shapiro.test(quail$hatch.mass)
shapiro.test(quail$hatch.mass[quail$sex=="Female"])
shapiro.test(quail$hatch.mass[quail$sex=="Male"])
```

```{r t Test}
t.test(hatch.mass ~ sex, data = quail)
```

This shows us that hatch mass does depend on the sex of the chick.

After exploring the dataset, formulate some additional tentative hypotheses that you wish to test.
I have provided some examples of the types of graphs that you could consider. 
You should alter these examples for your own use (and fix any issues that find.)

```{r Data Visualization}

ggplot(quail) +
  aes(x = hatch.mass) +
  geom_histogram(bins=100) +
  theme_cowplot()

ggplot(quail) +
  aes(x = hatch.mass, color = sex, fill = sex) +
  geom_histogram(binwidth = 0.5, position = "dodge") +
  theme_cowplot()

ggplot(quail) +
  aes(x = hatch.mass,  fill = sex) + 
  geom_density(alpha=.3) +
  theme_cowplot()
```

There appears to be some other factors influencing the hatch mass, in addition to sex. 
Explore the data and determine how to test a hypothesis about one of these other factors.

-------------------------------------------------------------------------------------------------------------------
New hypothesis: The investment of the parents influences chick mass.

The data I want to explore in more detail uses the quantitative variables egg mass, hatch mass, and week2 mass.The 
hatch mass variable has already been visualized so only egg mass and week2 mass are shown. The simple eda shows 
both variables appear to be normal.
```{r Data exploration}
simple.eda(quail$`egg.mass`)
summary(quail$`egg.mass`)
simple.eda(quail$week2.mass)
summary(quail$week2.mass)
```

Housekeeping for the categorical variables of interest. Recoded maternal and paternal lines as well as created a 
new variable called "parents" which identifies parental investment crosses. Changed hatching day to factor.

There are some NA values in the data which are removed for later visualizations.

https://www.rdocumentation.org/packages/photobiology/versions/0.9.30/topics/na.omit
Exploratory data analysis in R. Ch. 4.9 mutate()
https://stackoverflow.com/questions/7201341/how-can-two-strings-be-concatenated
https://www.math.ucla.edu/~anderson/rw1001/library/base/html/paste.html
https://stackoverflow.com/questions/36615769/is-there-an-alternative-to-revalue-function-from-plyr-when-using-dplyr
```{r Housekeeping}
quail <- na.omit(quail) #remove NA

m_line <- recode(quail$maternal.line, "H" = "Maternal high", "L" = "Maternal low")
p_line <- recode(quail$paternal.line, "H" = "Paternal high", "L" = "Paternal low")
parental_line = data.frame(m_line, p_line)
summary(parental_line)

quail <- mutate(quail, parents = paste(m_line, p_line, sep=" x "))
table(quail$parents)

quail$hatching.day <- as.factor(quail$hatching.day)
table(quail$hatching.day)
```

Perhaps staying in the egg for an additional day allows a chick to grow more, or maybe it doesn't.
We can't run a simple eda because the variable is no longer numeric but the table from before 
tells us hatching day is very much weighted towards day 17. Taking a quick peek with a boxplot 
reveals suggests there wasn't much difference in hatch mass based on what day the chick was born. 
Visualizing the means also doesn't suggest a promising data exploration. So we're gonna scrap 
that.
```{r Quick look at hatch day and hatch mass}
ggplot(quail) + 
  aes(x = hatching.day, y = hatch.mass) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("Hatch day") + 
  ylab("Hatch mass (g)") +
  theme_cowplot()

plotmeans(hatch.mass ~ hatching.day, data = quail, xlab = "Hatch day", ylab = "Hatch mass (g)")
```


To check the data that was selected shapiro tests were carried out. I thought using logic and for loops would make 
the process easier but it got so gross so fast. 

Looking at the results of the shapiro tests we can see several variables yielding p-values less than 0.05. The skew
on these variables is for the most part between -0.5 and 0.5 which allows us to say the distribution is 
apporximately symmetrical. Looking at the distribution of data with the simple eda we can see the data isn't 
horrible. Both variables are right skewed and between -0.5 and 0.5; therefore we say the two are approximately 
symmetrical (according to Brownmath.com) so we can probably ignore it. The greatest skew comes from looking at 
Maternal low x Paternal high (hatch mass) which had p-value: 0.00607270038906764 and skew: -0.523814340748249. 
Even so the skew is small enough that we can almost say it is approximately symmetrical.

Distribution of the parental line data is well balanced

p-value report:
hatch mass

 Maternal high p-value:  0.017188649785972
 Maternal low p-value:  0.124122738179463
 Paternal high p-value:  0.122927108707938
 Paternal low p-value:  0.00172976930963811
 Maternal high x Paternal low p-value:  0.22400832115948
 Maternal low x Paternal high p-value:  0.00607270038906764
 Maternal low x Paternal low p-value:  0.0730605593584457

egg mass

 Maternal high p-value:  0.163379596841482
 Maternal low p-value:  0.0591312769918625
 Paternal high p-value:  0.141950593175119
 Paternal low p-value:  0.00316749563148435
 Maternal high x Paternal high p-value:  0.24063069078182
 Maternal high x Paternal low p-value:  0.190064322294475
 Maternal low x Paternal high p-value:  0.0138178345042709
 Maternal low x Paternal low p-value:  0.200543219527835

https://brownmath.com/stat/shape.htm
https://www.datamentor.io/r-programming/for-loop/
https://www.datamentor.io/r-programming/if-else-statement/
https://stackoverflow.com/questions/9317830/r-do-i-need-to-add-explicit-new-line-character-with-print/9317914
```{r Checking out the variables}
m_p <- c("Maternal", "Paternal")
l_h <- c("high", "low")
parentz <- c("Maternal high x Paternal high", "Maternal high x Paternal low", "Maternal low x Paternal high", "Maternal low x Paternal low")

cat("hatch mass\n")
for (a in m_p){
  for (b in l_h){
    name = paste(a, b)
    if (a == "Maternal") 
      {shap = shapiro.test(quail$hatch.mass[m_line==name])$p.value 
      cat(paste("\n", name, "p-value: ", shap))
      if (shap < 0.05) 
        {cat(paste("\n", "    ^^ skew: ", skewness(quail$hatch.mass[m_line==name])))}}
    else if (a == "Paternal") {shap = shapiro.test(quail$hatch.mass[p_line==name])$p.value
    cat(paste("\n", name, "p-value: ", shap))
      if (shap < 0.05) 
        {cat(paste("\n", "   ^^ skew: ", skewness(quail$hatch.mass[p_line==name])))}}
    }
  }

for (val in parentz){
  shap = shapiro.test(quail$hatch.mass[quail$parents==val])$p.value
  cat(paste("\n",val, "p-value: ", shap))
  if (shap < 0.05) 
    {cat(paste("\n    ^^ skew: ", skewness(quail$hatch.mass[quail$parents==val])))}
  }

cat("\n\negg mass\n")
for (a in m_p){
  for (b in l_h){
    name = paste(a, b)
    if (a == "Maternal") 
      {shap = shapiro.test(quail$egg.mass[m_line==name])$p.value 
      cat(paste("\n", name, "p-value: ", shap))
      if (shap < 0.05) 
        {cat(paste("\n", "    ^^ skew: ", skewness(quail$egg.mass[m_line==name])))}}
    else if (a == "Paternal") {shap = shapiro.test(quail$egg.mass[p_line==name])$p.value
    cat(paste("\n", name, "p-value: ", shap))
      if (shap < 0.05) 
        {cat(paste("\n", "   ^^ skew: ", skewness(quail$egg.mass[p_line==name])))}}
    }
  }

for (val in parentz){
  shap = shapiro.test(quail$egg.mass[quail$parents==val])$p.value
  cat(paste("\n",val, "p-value: ", shap))
  if (shap < 0.05) 
  {cat(paste("\n    ^^ skew: ", skewness(quail$egg.mass[quail$parents==val])))}
    }

cat("\n\n")
table(m_line)
table(p_line)
table(quail$parents)
```

Boxplots were created to better see the individual effects of parental lines on hatch mass. We can see from the 
boxplots and plotted means that maternal investment appears to contribute the most to hatch mass. The difference in
mean hatch mass is negligible when we look at the paternal line. Similar observations were made for the egg mass.
```{r Boxplots and mean comparison of hatch mass by parental line}
ggplot(quail) + 
  aes(x = m_line, y = hatch.mass) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "red") +
  guides(alpha=FALSE) +
  xlab("Maternal line") + 
  ylab("Hatch mass (g)") +
  theme_cowplot()

ggplot(quail) + 
  aes(x = p_line, y = hatch.mass) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("Paternal line") + 
  ylab("Hatch mass (g)") +
  theme_cowplot()

plotmeans(hatch.mass ~ m_line, data = quail, xlab = "Maternal line", ylab = "Hatch mass (g)")
plotmeans(hatch.mass ~ p_line, data = quail, xlab = "Paternal line", ylab = "Hatch mass (g)")

ggplot(quail) + 
  aes(x = m_line, y = egg.mass) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "red") +
  guides(alpha=FALSE) +
  xlab("Maternal line") + 
  ylab("Egg mass (g)") +
  theme_cowplot()

ggplot(quail) + 
  aes(x = p_line, y = egg.mass) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("Paternal line") + 
  ylab("Egg mass (g)") +
  theme_cowplot()

plotmeans(egg.mass ~ m_line, data = quail, xlab = "Maternal line", ylab = "Egg mass (g)")
plotmeans(egg.mass ~ p_line, data = quail, xlab = "Paternal line", ylab = "Egg mass (g)")
```

T-tests were used to test the significance of the difference in means between high and low parental investments. 
Unsurprisingly the the difference between high and low investment in the maternal lines was significant for both 
hatch mass and egg mass (p = 2.2e-16 and 2.2e-16 respectively). The difference between paternal lines was not 
significant for both hatch mass and egg mass (p = 0.3083 and 0.229 respectively). 
```{r t-test}
t.test(hatch.mass ~ m_line, data = quail)
t.test(hatch.mass ~ p_line, data = quail)
t.test(egg.mass ~ m_line, data = quail)
t.test(egg.mass ~ p_line, data = quail)
```

Boxplots and plots of means were used to visualize the combined effects of parental lines (maternal and paternal). 
I really thought this would be a coolthing to visualize but it's just repeating or reaffirming the last set of 
boxplots. Anyways, maternal investment is suggested to be a dominating factor in both egg mass and hatch mass.

```{r Boxplots and mean comparison of parental investment by parental crosses}
ggplot(quail) + 
  aes(x = parents, y = hatch.mass) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("Parent cross") + 
  ylab("Hatch mass (g)") +
  theme_cowplot()

ggplot(quail) + 
  aes(x = parents, y = egg.mass) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("Parent cross") + 
  ylab("Egg mass (g)") +
  theme_cowplot()

plotmeans(hatch.mass ~ parents, data = quail, xlab = "Parents", ylab = "Hatch mass (g)")
plotmeans(egg.mass ~ parents, data = quail, xlab = "Parents", ylab = "Egg mass (g)")
```

ANOVA confirms the conclusions made from the boxplots and plots of means. There are significant differences in the
means of different populations. With p-values < 0.05 for hatch mass and egg mass (p = 2e-16 for both). 
Going further with the Tukey test it's made even more clear that maternal factors have the greatest influence on
early mass. 

Hatch mass:Maternal high x Paternal low-Maternal high x Paternal high p = 0.4050669, 
Maternal low x Paternal low-Maternal low x Paternal high p = 0.9911414. 
Egg mass: Maternal high x Paternal low-Maternal high x Paternal high p = 0.2087676,
Maternal low x Paternal low-Maternal low x Paternal high p = 0.9936364.

The high p-values indicate statistically insignificant differences between the populations where the Maternal line 
is the same. 

http://www.sthda.com/english/wiki/one-way-anova-test-in-r
```{r}
cat("Hatch mass")
hatch.aov <- aov(hatch.mass ~ parents, data = quail)
summary(hatch.aov)
TukeyHSD(hatch.aov)

cat("\nEgg mass")

egg.aov <- aov(egg.mass ~ parents, data = quail)
summary(egg.aov)
TukeyHSD(egg.aov)
```

Originally I was trying to help Fallon figure out a way to visualize multiple boxplots of different y-values in a 
single figure. This was not what she wanted but it looks cool so it stays. The egg and hatch mass boxplots are 
nothing new but looking at the week 2 mass suggests regardless of the chick's parentage, they seem to approach a 
similar mass. In other words, it may be that parental investment has the most impact around hatching. As time goes 
on the effect of parental investment on mass is less pronounced.

https://stackoverflow.com/questions/14785530/ggplot-boxplot-of-multiple-column-values
https://stackoverflow.com/questions/14604439/plot-multiple-boxplot-in-one-graph
https://www.rdocumentation.org/packages/mudata/versions/0.1.1/topics/id.vars
```{r Boxplot of egg, hatch, and week 2 mass}
mass <- melt(quail, id.vars='parents', measure.vars= c('egg.mass','hatch.mass', 'week2.mass'))
ggplot(mass) +
      geom_boxplot(aes(x = parents, y = log(value), color = variable)) + 
      xlab("Parent cross") +
      ylab("log10 mass (g)") +
      theme_cowplot()
```

Egg and hatch mass are very likely dependent variables but we can use the two to see the relationship between chick
mass and parental investment in another way. From the scatterplot we can confirm again that the individuals with
the greatest egg and hatch mass are those with high maternal investment. 
```{r Scatterplot egg mass x hatch mass}
ggplot(quail) + 
  aes(y = (hatch.mass), x = (egg.mass)) + 
  theme(legend.position = c(0.05, 0.8)) +
  geom_point(aes(colour = parents), alpha = 0.5) +
  geom_smooth(method=lm , colour="black", se=TRUE) +
  ylab("Hatch mass (g)") + 
  xlab("Egg mass (g)") +
  theme_cowplot()
```
The p-value for linear regression is well below 0.05 (p = 2.2e-16). 
```{r Statistics on linear regression}
egg_mass_fit = lm((egg.mass)~(hatch.mass), data=quail)
summary(egg_mass_fit)
```

Out of curiosity I wondered if the egg itself was something that also differed in mass. To investigate this 
question I created a new variable called just_the_egg which was the result of subtracting the hatch mass from the 
egg mass. The variable was plotted as a boxplot and plot of means with respect to parental line cross and suggests that maternal investment is the deciding factor for just the egg as well.
```{r Just the egg}
quail <- mutate(quail, just_the_egg = (egg.mass - hatch.mass))

ggplot(quail) + 
  aes(x = parents, y = just_the_egg) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("Parental cross") + 
  ylab("just_the_egg mass (g)") +
  theme_cowplot()

plotmeans(hatch.mass ~ parents, data = quail, xlab = "Parents", ylab = "just_the_egg mass (g)")
```


https://twitter.com/CMastication/status/1142832452890759169/photo/1