---
title: "Homework 05"
author: "Rob Han"
date: "2/25/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
Use the following syntax to load any libraries that you need. This code will prevent errors that will interfere with knitting of the Rmd file.
Also note that the `include=FALSE` option will prevent this code chunk from appearing in the markdown file.

We will load the entire tidyverse library, which includes *dplyr*, *ggplot2*, *readr*, and other packages.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("UsingR")) install.packages("UsingR"); library(UsingR)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("e1071")) install.packages("e1071"); library(e1071)  

```

I recently found out that competitive pokemon battling is kind of a thing which is pretty cool and brings me back to when I was
obsessed with pokemon in middle-school. When I was a kid I really wanted to be a gym leader... A pokemon gym leader. Many 
pokemon gym leaders will follow a theme when it comes to picking their team members. So I wonder, can data analysis help me 
choose what type my gym should be?

Main Database With Generation and Legendary flag.

Accessed from *https://www.kaggle.com/abcsds/pokemon*

## Data import and checking

The datafile itself is in csv format and is of a decent size.
```{r Import data}
project_data <- read.csv("Pokemon.csv")

```

Check the data, make sure everything has been imported. 
```{r Check data}
str(project_data)
cat("rows", nrow(project_data))
cat("\ncolumns", ncol(project_data))

head(project_data)
tail(project_data)
```
Checking out the dataset we can see that we have a number of variables which describe some basic information about each 
Pokemon, notably their battle stats and Legendary status. According to the Pokemon company there are currently 807 Pokemon.
Since this dataset only includes 721 Pokemon up to generation 6, it's likely that the dataset is missing Pokedex entries from
the most recent games. This is okay because I stopped playing Pokemon after generation 4 so I could care less about these
new-fangled pokemons. We don't immediately see any NA's however that doesn't mean they aren't present. We'll check that next.

To check if we have NA's in the dataset that aren't found by head/tail sampling we can use the following code to count NA 
instances.
```{r Count NA}
sum(is.na(project_data)) 
```
Since the number of NA's is 0 we will not need to take care and account for them later when making graphs.

## Data wrangling

At this stage we can proceed to cut down the dataset to suit our purposes. Since I have no a mild attatchment to the pokemon
I used when I was a kid we'll filter out the later generations of pokemon.
```{r Filter out the uncool pokemon}
pokemon <- dplyr::filter(project_data, Generation <= 4)
```

In pokemon battles, being able to make the first move can often dictate the flow of the entire battle. Some people will opt for
pokemon that can debuff the oponents team, but lets be honest overwhelming power is the way to go. Our opponent needs to 
be utterly humiliated for full immersion; Ash Ketchum never used a status debuff (to my knowledge) so neither will I. What I
want is an opener that can hit hard and fast so we'll look specificaly at the attack, special attack, and speed stats of 
various pokemon.

To check the distribution of these stats we can visualie with a simple eda and run the Shapiro normality test.
```{r EDA of selected stats}
simple.eda(pokemon$Attack)
simple.eda(pokemon$Sp..Atk)
simple.eda(pokemon$Speed)

shapiro.test(pokemon$Attack)
shapiro.test(pokemon$Sp..Atk)
shapiro.test(pokemon$Speed)

skewness(pokemon$Attack)
skewness(pokemon$Sp..Atk)
skewness(pokemon$Speed)
```
The Shapiro test returned results indicating that the hypothesis that the data is normally distributed should be rejected
(Attack: 1.466e-07, Special Attack: 8.404e-11, Speed: 4.799e-06). Taking a look at the skew we can see that all of our 
variables are right skewed (Attack: 0.5721488, Special Attack: 0.7410125, Speed: 0.3936992). There are some things we could 
try right away to get a more normal distribution but we'll leave it be for now.

```{r Normies vs legendaries}
ggplot(pokemon, aes(x = Speed, y = Attack, color = Legendary)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()
ggplot(pokemon, aes(x = Speed, y = Sp..Atk, color = Legendary)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()
ggplot(pokemon, aes(x = Attack, y = Sp..Atk, color = Legendary)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()
```

The scatterplot with glm overlay shows legendary pokemon are not only some of the fastest pokemon, they also appear to be some 
of the strongest in terms of attack/special attack power. These legendary pokemon are definitely the ones contributing the most
to the skew in our dataset.

That being said legendary pokemon aren't exactly the easiest to come by. It might be more useful to look at the same
relationship while excluding legendary pokemon. In this scenario we might distinguish pokemon by their primary type. Since
there are a lot of types we should seperate into two or three subsets.
https://stackoverflow.com/questions/5584597/how-to-filter-a-tables-row-based-on-an-external-list/53416497
```{r Group by type}
bug_to_fighting <- c("Bug", "Dark", "Dragon", "Electric", "Fairy", "Fighting") 
fire_to_normal <- c("Fire", "Ghost", "Grass", "Ground", "Ice", "Normal")
posion_to_water <- c("Poison", "Psychic", "Rock", "Steel", "Water")
# Filter out legendary pokemon
non_legendary <- dplyr::filter(pokemon, Legendary == "False") 
# Create subset df's
Bug_Grass <- subset(non_legendary, Type.1 %in% bug_to_fighting)
Fight_Ground <- subset(non_legendary, Type.1 %in% fire_to_normal)
Posion_Water <- subset(non_legendary, Type.1 %in% posion_to_water)

#Recheck population distribution
shapiro.test(pokemon$Attack)
shapiro.test(pokemon$Sp..Atk)
shapiro.test(pokemon$Speed)

skewness(pokemon$Attack)
skewness(pokemon$Sp..Atk)
skewness(pokemon$Speed)
```

After filtering out the legendary pokemon the Shaprio test still does not give us p-values we'd like to see (Attack: 1.782e-06,
Special Attack: 1.124e-08, Speed: 1.674e-05). Checking out the skewness (Attack: 0.5268547, Special Attack: 0.6303942, Speed:
0.3784766), the right skew seems to have been tamed a little but has not reached a level where we can say the skew is
neglibible. Although we have removed the legendary pokemon we haven't removed Mega-forms which were added after generation IV.
These pokemon are not counted as legendaries and are a suped up form of a normal pokemon with legendary-level stats. Right now
we can't remove these pokemon however they will be removed later.

```{r Evaluate speed vs Atk/Special Attack}
ggplot(Bug_Grass, aes(x = Speed, y = Sp..Atk, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Special Attack") +
  theme_cowplot()
ggplot(Bug_Grass, aes(x = Speed, y = Attack, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Attack") +
  theme_cowplot()

ggplot(Fight_Ground, aes(x = Speed, y = Sp..Atk, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Special Attack") +
  theme_cowplot()
ggplot(Fight_Ground, aes(x = Speed, y = Attack, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Attack") +
  theme_cowplot()

ggplot(Posion_Water, aes(x = Speed, y = Sp..Atk, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Special Attack") +
  theme_cowplot()
ggplot(Posion_Water, aes(x = Speed, y = Attack, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Attack") +
  theme_cowplot()
```
When we break it down this way we can get a general sense for what pokemon type has the highest performers are with regard to
my needs for an opener. We'll pick a few of the best looking types and compare those together

```{r}
best1 <- c("Psychic", "Ghost", "Dragon", "Grass")
high_performers <- subset(non_legendary, Type.1 %in% best1)

ggplot(high_performers, aes(x = Speed, y = Sp..Atk, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Special Attack") +
  theme_cowplot()
ggplot(high_performers, aes(x = Speed, y = Attack, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Attack") +
  theme_cowplot()
ggplot(high_performers, aes(x = Attack, y = Sp..Atk, color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  ggtitle("Damage") +
  theme_cowplot()
```

Comparing stats is great and all but how can we be sure that they will tell the whole story. Obviously there are a lot of other
factors which can influence the outcome of a battle. Therefore, it would be very nice if we had some pokemon battle data... Oh
wait we do. The Weedle's cave challenge was originally intended as a machine learning project to predict the outcome of a 1v1
pokemon battle using a large dataset of random 1v1 pokemon battles. Unfortunately the algorithim determining the dataset's
winners was unable to take into account many factors such as pokemon moves, and trainer decisions; however since the simulation
can't use those factors it must rely purely on pokemon stats which is exactly what we've been looking at. Lets load in the data
and see what we have.

Using data from the Weedle's cave challenge we can estimate the winrate ratio of each pokemon.
https://stackoverflow.com/questions/5577727/is-there-an-r-function-for-finding-the-index-of-an-element-in-a-vector

```{r Integrate Weedles cave challenge data}
# Load data
battle_data <- read.csv("combats.csv")
# Filter out later generation pokemon
battle_data <- dplyr::filter(battle_data, Winner < 493)
battle_data <- dplyr::filter(battle_data, First_pokemon < 493)
battle_data <- dplyr::filter(battle_data, Second_pokemon < 493)
# Initialize column variables in pokemon df
pokemon$Wins <- 0
pokemon$Battles <- 0

# Populate new column variables in pokemon df
for (i in 1:length(battle_data$Winner)){
  # Battles
  pmon_index1 <- match(battle_data$First_pokemon[i], pokemon$X.) # Find index of pokemon in pokemon df using pokedex no.
  pmon_index2 <- match(battle_data$Second_pokemon[i], pokemon$X.)
  pmon_index1 <- pmon_index1[1] # In some cases there are two pokedex entries for the same pokemon bc of "Mega Evolutions" idk its too new for me lol. Anyways this makes sure we only get the index of the original pokemon form.
  pmon_index2 <- pmon_index2[1]
  pokemon[pmon_index1, 15] <- pokemon[pmon_index1, 15] + 1 # Add to relevant count
  pokemon[pmon_index2, 15] <- pokemon[pmon_index2, 15] + 1
  # Wins
  pmon_index3 <- match(battle_data$Winner[i], pokemon$X.)
  pmon_index3 <- pmon_index1[1]
  pokemon[pmon_index3, 14] <- pokemon[pmon_index3, 14] + 1}

pokemon <- dplyr::filter(pokemon, Battles != 0) # Filter out those random "Mega Evolutions"
pokemon <- dplyr::filter(pokemon, Legendary == "False")
# Handles win rate calculations
pokemon$WR <- 0
for (i in 1:length(pokemon$WR)){
  pokemon$WR[i] = pokemon$Wins[i]/pokemon$Battles[i]
  }
```

We now have simulated winrate data from the Weedle's cave challenge associated with our original pokemon dataset. 
Furthermore, the mega forms have been removed from the data set. What does this winrate data look like?

```{r}
simple.eda(pokemon$WR)
shapiro.test(pokemon$WR)
skewness(pokemon$WR)
```
According to the Shapiro test win rate is likely normally distributed (p = 0.2418) with skewness = 0.09774497 so not very
skewed. Since we know that the winrate data looks pretty good why don't we check out what happens if we check our "high
performers" with the winrate data.
```{r WR data vs stats of "high performers"}

best1 <- c("Psychic", "Ghost", "Dragon", "Grass")
high_performers <- subset(pokemon, Type.1 %in% best1)

ggplot(high_performers, aes(x = Attack, y = (WR), color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()

ggplot(high_performers, aes(x = Sp..Atk, y = (WR), color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()

ggplot(high_performers, aes(x = Speed, y = (WR), color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()
```

Interestingly it seems that only dragon type pokemon benefit from increases in either speed, attack, or special attack
stats. Looking closer, we see that the most pronounced benefit seems to come from a higher speed stat.


To be honest, I never really saw myself as a Psychic, Ghost, or Grass gym leader. I really liked Water, Fire, Dragon, and
Fighting pokemon so lets run the same modeling with those choices instead.
```{r Investigate personal preferences}
choices <- c("Water", "Fire", "Dragon", "Fighting")
my_choices <- subset(pokemon, Type.1 %in% choices)

ggplot(my_choices, aes(x = Attack, y = (WR), color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()

ggplot(my_choices, aes(x = Sp..Atk, y = (WR), color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()

ggplot(my_choices, aes(x = Speed, y = (WR), color = Type.1)) +
  geom_point() +
  geom_smooth(method = "glm") +
  theme_cowplot()
```

I really like the way the fighting type looks out in terms of offensive power. Especially when looking at attack, the win rate
for fighting types seems to scale heaviliy with the attack and special attack stats. They are a little lacking speed-wise when 
compared to the other selected types but there are ways to get around going second.

My hypothesis: Hitting hard and and fast is in my opinion the best way to beat down all those pesky challengers to my 
hypothetical pokemon gym.

```{r GLM}
fighting <- subset(pokemon, Type.1 == "Fighting")
attack_glm <- glm(WR ~ Attack, data = fighting)
sp_attack_glm <- glm(WR ~ Sp..Atk, data = fighting)
speed_glm <- glm(WR ~ Speed, data = fighting)

cat("Attack:", attack_glm$aic)
cat("\nSpecial Attack:", sp_attack_glm$aic)
cat("\nSpeed:", speed_glm$aic)
```
The AIC results for the specific attributes I've chosen are not hugely different from one another (Attack: -27.48543, 
Sp.Atack: -24.9825, Speed: -24.53929). The attack stat is 3 points lower than the others indicating a superior goodness of fit.

```{r full GLM}
fighting_fullglm = glm(WR ~ Attack + Sp..Atk + Defense + Sp..Def + Speed + HP, data = fighting)
summary(fighting_fullglm)
```
In the full GLM we actually see several variables with pretty low Pr values but not quite what we want to see. We can use 
the dropterm function to see if we can improve our AIC. The objective of using dropterm is removing superfluous predictors.

```{r}
if (!require("MASS")) install.packages("MASS"); library(MASS)
dropterm(fighting_fullglm)
```

Lets try dropping the HP variable.
```{r}
fighting_fullglm2 = glm(WR ~ Attack + Sp..Atk + Defense + Sp..Def + Speed, data = fighting)
dropterm(fighting_fullglm2)
```

```{r}
fighting_fullglm3 = glm(WR ~ Attack + Sp..Atk + Defense + Sp..Def, data = fighting)
dropterm(fighting_fullglm3)
```

You may have noticed that the difference in AIC between each term isn't that large. This isn't a surprising observation
since the outcomes of the simulated pokemon battles were likely based on a calculation using pokemon stats. 

```{r}
fighting_fullglm4 <- glm(WR ~ Attack + Defense + Sp..Def, data = fighting)
dropterm((fighting_fullglm4))
```

Surprisingly (to me at least) it seems like the top three predictors for win rate with a fighting type pokemon are their
attack, defense and special defense stats.