---
title: "Homework 04"
author: "Biology Student"
date: "2/25/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
Use the following syntax to load any libraries that you need. This code will prevent errors that will interfere with knitting of the Rmd file.
Also note that the `include=FALSE` option will prevent this code chunk from appearing in the markdown file.

We will load the entire tidyverse library, which includes *dplyr*, *ggplot2*, *readr*, and other packages.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
```

## Objectives

For this assignment you will need to identify a dataset from a public data archive use it to conduct the following steps:
- [ ] Data Wrangling
- [ ] Data Exploration
- [ ] Data Visualization
- [ ] Hypothesis testing using a generalized linear model

Your first step is to identify the dataset that you want to use and confirm that it is suitable for this assignment.
1. Contains between 100 and 10,000 observations.
2. Has a quantitative outcome variable.
3. Measures 4 or more potential explanatory variables. At least one quantitative variable.

## Public Data Archives

You can find your dataset either by browsing a data archive or by finding a paper that includes the raw data as supplementary material (or links to a data archive).
There are many specialized data archives (http://oad.simmons.edu/oadwiki/Data_repositories), but these are some general ones that include biology:

- Dryad: http://www.datadryad.org/
- Figshare: https://figshare.com/
- Mendeley Data: https://data.mendeley.com/
- TreeBASE: https://www.treebase.org
- Zenodo: https://zenodo.org/
- Dataverse: https://dataverse.harvard.edu/

