---
title: "Homework 03"
author: "Robert Han"
date: "02/03/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 3
Due at noon on 12 Feb 2020.

## Loading Libraries
Use the following syntax to load any libraries that you need. This code will prevent errors that will interfere with knitting of the Rmd file.
Also note that the `include=FALSE` option will prevent this code chunk from appearing in the markdown file.

First we will load the "conflicted" library. This will help us identify any problems with packages that we use.

We will then load the entire tidyverse library, which includes *dplyr*, *ggplot2*, *readr*, and other packages.

```{r Load Libraries, include=FALSE}
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
# Load other packages here.
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
```

# Exploratory Data Analysis Checklist

Follow Chapter 5 from *Exploratory Data Analysis with R* (https://leanpub.com/exdata/read_full) by Roger D. Peng to explore one of the following datasets:
- Uganda Bat Dataset from Lab 4
- NHANES Dataset from Lab 3


Loading in the data is the first step. The Uganda Bat Dataset is loaded in from a .csv file in the working
directory. A line of code is also run so that the date format used in the data is correctly represented in R.
```{r Load Data}
if (!require("NHANES")) install.packages("NHANES"); library(NHANES)
data("NHANES")
BatData <- read_csv("UgandaBatsFilteredMetrics.csv", 
    col_types = cols(`Collection date` = col_date(format = "%m/%d/%y")))
```


Running the summary function on the data we see we have 103 observations, 5 categorical variables (Collection.data, Season, Sex, Class, and Habitat description), and 6 continuous variables (Elevation, Total.Length, Hind.Foot, Ear, FA.length, and Mass).
```{r Summary of Bat Data, echo = FALSE}
str(BatData)
```

```{r}
ggplot(BatData) +
  aes(x = Elevation, y = Mass, color = Sex) +
  geom_point() +
  theme_cowplot()



```


Remember to edit the Rmd to remove any irrelevant code for your assignment.

## Formulate your question

Always formulate your question(s) *before* you begin analyzing the data. 
(But you may look at the names of the variables measured to see what questions are possible.)

## Read in your data

## Check the packaging

Validate that the data file is complete and properly loaded.

## Run str()

Fix any problems that you find.

## Look at the top and the bottom of your data

## Check your “n”s

## Validate with at least one external data source

## Try the easy solution first

## Challenge your solution

## Follow up

# Analytic Graphics

Follow Section 6.1 and 6.2 to visualize the comparisons that were essential to answering the question.

## Show comparisons

## Show causality, mechanism, explanation, systematic structure
