---
title: "Homework 03"
author: "Robert Han"
date: "02/03/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 3
Due at noon on 12 Feb 2020.

## Loading Libraries
```{r Load Libraries, include=FALSE}
if (!require("conflicted")) install.packages("conflicted"); library(conflicted)
# Load other packages here.
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("NHANES")) install.packages("NHANES"); library(NHANES)
if (!require("gplots")) install.packages("gplots"); library(gplots) 
if (!require("corrplot")) install.packages("corrplot"); library(corrplot)
```

# Exploratory Data Analysis Checklist

Follow Chapter 5 from *Exploratory Data Analysis with R* (https://leanpub.com/exdata/read_full) by Roger D. Peng to explore one of the following datasets:
- Uganda Bat Dataset from Lab 4
- NHANES Dataset from Lab 3


Loading in the data is the first step. "This is survey data collected by the US National Center for Health 
Statistics (NCHS) which has conducted a series of health and nutrition surveys since the early 1960’s. 
Since 1999 approximately 5,000 individuals of all ages are interviewed in their homes every year and complete the 
health examination component of the survey. The health examination is conducted in a mobile examination
center (MEC)" (Lab03 description). This dataset includes data from two years (2009-10 & 2010-11)

According to what's been read in, we can expect 10,000 rows of data, with several rows having at least one NA value, and 76 columns or variables.
```{r NHANES}
data("NHANES")
str(NHANES)
#help(NHANES) #information
nrow(NHANES)
ncol(NHANES)
```

Summary of dataset.
```{r Summary of NHANES dataset, include=FALSE}
summary(NHANES)
```

Looking at the head and tail of the dataset we see several NA's indicating we may need to run an NA removal
function at some point. 
```{r Peeksie at the data}
head(NHANES)
tail(NHANES)
```

## Formulate your question

Question: Recently I came across a subreddit called r/EOOD, or exercise out of depression. We've all heard of the 
"runners high" but can physical activity really reduce depression?

## Validate with at least one external data source

We can calculate the number of respondents reporting some form of depression using the Depressed variable in NHANES. 
Out of the participants choosing to answer 21.4% reported some level of depression. My cursory look into depression
statistics revealed major depressive episodes are the common mode of reporting, and that the NIH estimates around 7%
of adults have at least one major depressive episode per year. If we say that answering "Most" to the Depressed
prompt, most closely resembles a major depressive episode we find that 6.3% of respondents fall into that analog. 
Therefore, it is probably safe to say this dataset is a fair representation.

https://www.nimh.nih.gov/health/statistics/major-depression.shtml
```{r}
cat("Depressed\n")
summary(NHANES$Depressed)
cat("Depressed:", 100*(1009+418)/(5246+1009+418), "%\n")
cat("Depressed (Most):", 100*(418)/(5246+1009+418), "%")
```

## Try the easy solution first
```{r Get subset of NHANES, include = FALSE}
NHANES_adult <- subset(NHANES, (Age >21)) #subset the full dataset
NHANES_adult <- drop_na(NHANES_adult, LittleInterest) 
NHANES_adult <- drop_na(NHANES_adult, Depressed) 
NHANES_adult <- drop_na(NHANES_adult, PhysActive) #NA removal
NHANES_adult <- drop_na(NHANES_adult, Testosterone) 
NHANES_adult <- drop_na(NHANES_adult, SleepHrsNight) 
```

To visualize the relationship between physical activity and depression, we can take a look at the Depressed and
PhysActive variables. Both variables are categorical so we'll take a look at counts in a barchart.

It kind of looks like the ratio of "No" to "Yes" (Physically Active) increases with the Depressed variable. Its tough
to tell because there is not a even distribution of counts between the levels of Depressed. To get a better look we
can use the same data to build a percent stacked barchart. Checking that out, we can see the percentage of people that
are not physically active is greatest in the depressed most days category, and the opposite is true for the depressed
"None" days.
```{r EZ comparison}
ggplot(NHANES_adult) +
  aes(x = Depressed, fill = PhysActive, na.rm = TRUE) + 
  geom_bar(position = "dodge") +
  theme_cowplot()

ggplot(NHANES_adult) +
  aes(x = Depressed, fill = PhysActive, na.rm = TRUE) + 
  geom_bar(position = "fill") +
  ylab("Percent") +
  theme_cowplot()
```

Seeing as little interest can be a symptom of depression we can do the same visualizations for the variable
LittleINterest and PhysActive. Unsurprisingly we get very similar barcharts for these two variables as the last.
```{r EZ comparison pt. 2}
ggplot(NHANES_adult) +
  aes(x = LittleInterest, fill = PhysActive, na.rm = TRUE) + 
  geom_bar(position = "dodge") +
  theme_cowplot()

ggplot(NHANES_adult) +
  aes(x = LittleInterest, fill = PhysActive, na.rm = TRUE) + 
  geom_bar(position = "fill") +
  ylab("Percent") +
  theme_cowplot()
```

We run a chi squared test on the two variables to confirm our thoughts about the relationship between being
physically active. We get a significant p-value (p = 1.787e-09) which means it is safe to reject the null
hypothesis that there is no relationship between the variables of interest. We can actually look at the
contribution of each cell in the table by checking out the Pearson residual for each. From that calculation we can
see that the combination of Physically Active: No and Depressed: Most days contributed the most to the chi squared
(r = 3.605).

http://www.sthda.com/english/wiki/chi-square-test-of-independence-in-r
https://stackoverflow.com/questions/48918408/how-to-add-label-in-table-in-r
```{r Chi-squared test}
PhysActivity_Depression = table(NHANES_adult$PhysActive, NHANES_adult$Depressed, dnn = c("Physically Active", "Depressed (days)"))
PhysActivity_Depression

chisq <- chisq.test(PhysActivity_Depression)
chisq 
round(chisq$residuals, 3)
```

## Challenge your solution
I had a little difficulty mustering up the creative juices to challenge my solution creatively so I decided to just
copy the method used in the Exploratory Data Analysis book. Even so, I understand why this step is an important one
in the data analysis process, since we took "the easy way out" earlier. To challenge the solution, I employed the 
bootstrap method which randomly alters the dataset in a way that it is for the most part the same but slightly 
different. The same statistical analysis was then used to analyze the "new" dataset. 

After bootstrapping, there is still a significant p-value (p = 6.565e-09) which is still statistically significant.
Furthermore, the Pearson residuals for several variable combinations decreased indicating that they had
less contribution to the overall chi squared value. Regardless, the Pearson residual for the combination of
Physically Active: No and Depressed: Most days contributed the most to the chi squared (r = 3.326).
```{r Challenge the solution}
set.seed(11)
N <- nrow(NHANES_adult)
idx <- sample(N, N, replace = TRUE)
NHANES_adult2 <- NHANES_adult[idx, ]

PhysActivity_Depression2 = table(NHANES_adult2$PhysActive, NHANES_adult2$Depressed, dnn = c("Physically Active", "Depressed (days)"))
PhysActivity_Depression2

chisq2 <- chisq.test(PhysActivity_Depression2)
chisq2
round(chisq2$residuals, 3)
```

## Follow up

To follow up on these findings, I'd argue that the variables selected to answer the intial question could have been
better. I think PhysActiveDays would have been a stronger representation of exercise since whether or not a person
is physically active can be subjective. I don't believe looking at how these results could change after several
years is necessary but could be interesting. 

# Analytic Graphics

I have no idea what analytical graphics to use but here are some that the page on chi squared I refrenced earlier
used. These graphics show a contingency table for the variables PhysActive and Depressed. It's cool because the
size of the circle correlates to the number of instances in the dataset. The second shows the scale of the Pearson
residuals from chi squared with the size and color of the circles correlating to magnitude of the individual
residuals. The default color for the text labels is actually red and it took me a while to figure out that I
had to change it to black lol (why is the default red???).
```{r Analytical analysis using graphical graphics}
# Visualize contingency table
balloonplot(t(PhysActivity_Depression),
    main ="Physical Activity~Depression", xlab ="Depressed (days)", ylab="PhysActive", label.size = 0.75,
    label = TRUE, show.margins = FALSE)

# Visualize Pearson residuals
corrplot(tl.col = "black", chisq$residuals, is.cor = FALSE)
```

## Show comparisons

There isn't a whole lot we can do comparison-wise since the variables I chose cover both ends of the spectrum. Had
a chose perhaps PhysAcitveDays which measures the number of days in a week that a respondent is physically active,
I could have used PhysActive: No as the comparison.

## Show causality, mechanism, explanation, systematic structure

Low testosterone can be a contributer to the onset or worsening of depression in both males and females.
Furthermore, exercise can result in increased testosterone. Luckily, testosterone levels were measured in this
dataset and can be used to establish causality.

Of course after running the code I realized that the researchers did not get testosterone levels for female
participants... 

The boxplot shows, the medians and distributions of testosterone level for both PhysActive categories are
approximately the same with PhysActive: Yes having a slightly higher median. However the median testosterone value for
Depressed: None is greater than that of both Depressed: Several and Most which supports the proposed causality between
testosterone and depression. Since the median and IQR of of testosterone level by PhysActive does not show much
meaningful information we can check the means as well. 
```{r Boxplots for testosterone}
NHANES_adult_male <- subset(NHANES_adult, (Sex == "Male"))
NHANES_adult_female <- subset(NHANES_adult, (Sex == "Female"))

ggplot(NHANES_adult_male) + 
  aes(x = PhysActive, y = (Testosterone)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("PhysActive") + 
  ylab("Testosterone level") +
  theme_cowplot()

ggplot(NHANES_adult_male) + 
  aes(x = Depressed, y = (Testosterone)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("PhysActive") + 
  ylab("Testosterone level") +
  theme_cowplot()
#Female shows nothing
ggplot(NHANES_adult_female) + 
  aes(x = PhysActive, y = (Testosterone)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(alpha = 0.3, position = "jitter", colour = "red") +
  guides(alpha=FALSE) +
  xlab("Depressed") + 
  ylab("Testosterone level") +
  theme_cowplot()
```

When checking out the plotted means, a relatively small difference between PhysActive categories becomes apparent.
That being said the 95% confidence interval does overlap a little bit It's difficult to predict what a t-test might
yield for these variables. Taking a look at the plotted means for testosterone level and Depressed suggests
testosterone level decreases the more often a person is depressed. Once again there is overlap in the 95% confidence intervals so a statistical test could help clear things up.
```{r Means: Testosterone ~ PhysActive, Testosterone ~ Depressed}
plotmeans((Testosterone) ~ PhysActive, data = NHANES_adult, xlab = "PhysActive", ylab = "Testosterone level")
plotmeans((Testosterone) ~ Depressed, data = NHANES_adult, xlab = "Depressed", ylab = "Testosterone level")
```

Since the testosterone data can only add meaningful information to our narrative for males, I think it's best to
find another variable in the dataset that will encompass both males and females. Not getting enough sleep/having
troubled sleep can also result in depression regardless of gender. Furthermore, according to myself, physical
activity can really help with establishing and maintaining a healthy sleeping pattern. Therefore, we may be able to
draw some measure of causality from the sleep length/qualitymeasurements included in this dataset.

Analyzing the boxplots there does not appear to be an immediate difference in the distribution of hours slept with
regard to how individuals answered PhysActive. The boxplot for Depressed and hours slept shows relatively similar
distributions between categories of Depressed. Notably Depressed: Most has a lower median and upper/lower quartiles
than the other categories of Depressed.
```{r Boxplots for testosterone}
NHANES_adult <- drop_na(NHANES_adult, SleepHrsNight)

ggplot(NHANES_adult) + 
  aes(x = PhysActive, y = SleepHrsNight) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("PhysActive") + 
  ylab("Hourse slept") +
  theme_cowplot()

ggplot(NHANES_adult) + 
  aes(x = Depressed, y = SleepHrsNight) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("Depressed") + 
  ylab("Hours slept") +
  theme_cowplot()
```

Since the median and IQR for PhysActive and hours slept did not yield much meaningful information we can take a
look at the plotted means of these variables. In this case there is a obvious difference between the mean hours
slept for the two responses to PhysActive. Furthermore the 95% confidence intervals do not overlap and so we could
expect a statistically significant difference between the two populations after performing a t-test. Taking a look 
at the second graphic which looks at hours slept and the Depressed variable, we see a general decrease in the mean 
hours slept as the Depressed variable "increases". Confidence intervals for this plot are also pretty good so
statistical analysis should fit the narrative.
```{r Means: SleepHrsNight ~ PhysActive, SleepHrsNight ~ Depressed}
plotmeans(SleepHrsNight ~ PhysActive, data = NHANES_adult, xlab = "PhysActive", ylab = "Hours slept")
plotmeans(SleepHrsNight ~ Depressed, data = NHANES_adult, xlab = "Depressed", ylab = "Hours slept")
```

Well I guess we should do a t-test or two. The t-test for testosterone level did return significantly different
populations (p = 0.04514). Performing the test on hours slept with respect to PhysActive response returned a
significant likliehood of independent populations (p = 8.33e-06).
```{r A t-test or two}
cat("Testosterone level ~ PhysActive")
t.test(NHANES_adult$Testosterone ~ NHANES_adult$PhysActive, var.equal = TRUE)

cat("SleepHrsNight ~ PhysActive")
t.test(NHANES_adult$SleepHrsNight ~ NHANES_adult$PhysActive, var.equal = TRUE)
```
