---
title: "Homework 04"
author: "Robert Han"
date: "2/25/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
Use the following syntax to load any libraries that you need. This code will prevent errors that will interfere with knitting of the Rmd file.
Also note that the `include=FALSE` option will prevent this code chunk from appearing in the markdown file.

We will load the entire tidyverse library, which includes *dplyr*, *ggplot2*, *readr*, and other packages.

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("UsingR")) install.packages("UsingR"); library(UsingR)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)

```

Brenna recieved an article which was somewhat related to the paper I was assigned titled "A long journey toward
reproducible results." The article itself was about several groups of researchers who were unable to reproduce the
findings of the author's research group. To resolve this issue, they teamed up to standardize the way they set up 
and performed their experiments and shared the data with each other in an online dataset. I searched around for it, 
and found it, however Petrasheck was not listed as a contributer to the dataset. In my desperation I reached out the 
legend himself and lo and behold he responded. He sent him his data in a .raw image file but alas, no matter what I used to
open it, I could not unlock it's secrets. Photoshop, Gimp, online image converters, nothing could crack this nut. And so, I
resorted to recreating a figure from the Supplementary Information which holds summary statistics used to generate the
figures. 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC5762131/
https://www.scripps.edu/faculty/petrascheck/

Read in the data from the dataset that I made from the supplementary information pdf. Make the variables "Strain" and "n" 
unique factors so their arrangment won't change in the barplot.

https://www.r-bloggers.com/use-unique-instead-of-levels-to-find-the-possible-values-of-a-character-variable-in-r/
```{r read in data to project_data}
project_data <- read.csv("Summary statistics.csv")
project_data$Strain <- factor(project_data$Strain, levels=unique(project_data$Strain))

project_data$n <- sapply(project_data$n, as.character)
project_data$n[is.na(project_data$n)] <- " "

project_data$n <- factor(project_data$n, levels=unique(project_data$n))
project_data$n <- as.factor(project_data$n)
```

Check the data, make sure everything has been imported, make sure the changes we made earlier carried through.
```{r Check data}
str(project_data)
head(project_data)
tail(project_data)
```

So this was a joy to figure out. Luckily I knew to use stat = "identity" from project 1. In order to replicate the bar-graph
in figure d, we need to seperate the treatment groups, while also keeping strains together. To do so we use the position =
"dodge" arguement in conjunction with fill = Treatment. This however makes the x-axis the strain so we also need to
overwrite that. Because the data we are overwriting with is a discrete variable we use a specific call "scale_x_discrete()."
Since that makes the x-axis discrete we need to to specify in the geom_line calls (group = 1)
The rest of the additions are cosmetic such as horizontal lines, error bars, bar labels, and axis labels.

https://stackoverflow.com/questions/40211451/geom-text-how-to-position-the-text-on-bar-as-i-want
https://community.rstudio.com/t/error-aesthetics-must-be-either-length-1-or-the-same-as-the-data-2-fill/15579/3
http://www.sthda.com/english/wiki/ggplot2-error-bars-quick-start-guide-r-software-and-data-visualization
https://stackoverflow.com/questions/35095129/override-x-axis-ticks
https://stackoverflow.com/questions/31818264/no-line-in-plot-chart-despite-geom-line/31819922
```{r barplot}
ggplot(project_data) +
  geom_line(aes(y =  40, group = 1)) +
  geom_line(aes(y =  30, group = 1)) +
  geom_line(aes(y =  20, group = 1)) +
  aes(x = Strain, y = Mean.lifespan, fill = Treatment) +
  geom_bar(stat="identity", width = 0.7, position = position_dodge(width = 1)) +
  geom_text(aes(label= Strain), position = position_dodge(width = 0.9), vjust = -1) +
  geom_errorbar(aes(ymin=Mean.lifespan-SEM, ymax=Mean.lifespan+SEM), width=.2, position=position_dodge(.9)) +
  scale_y_continuous(limits = c(0,45), expand = c(0, 0)) +
  scale_x_discrete(labels = c("166  133", "80   ", "    ", "32  30", "24  32", "24  30", "32  32", "32  48", "31  42", "43  55", "31  40", "32  40", "48  64", "64  71", "32  30", "40  56")) +
  xlab("Number of wells assayed") +
  ylab("Mean lifespan (days)") +
  theme_cowplot() 

```

When comparing this recreated figure with the original, I noticed that the error bars differed greatly between the two. In 
my figure the error bars are significantly larger than those in the paper. Checking the supplementary data and figure 
methods, I'm not exactly sure why mine are so much larger. Ser-7 for example was reported as having SEM of 11 which is
clearly reflected in my plot but appears much smaller in the paper especially considering that the figure in the paper is
cropped. As far as reproducibility goes, I'd give this paper a lower score because the raw data is not readily available.
Furthermore, the supplementary information does not appear to completely reflect the figures especially where these error
bars are concerned.


## Acknowledgements

Brenna, Fallon, and I hung out to lament about our troubles with finding data, being able to wallow in a group is so much
better than doing it alone. Also big shoutout to Michael Petrascheck PhD, who not only responded to my calls for help in a
timely manner on a Sunday but was able to send me the data for his paper, albeit in a format that I cannot open. 