---
title: "Lab 05 Warmup"
author: "Prof. Ken Field, Bucknell University"
date: "2/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load Libraries, include=FALSE}
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
```

## Loading Data

Scientists have been asked to investigate whether jelly beans cause acne. 
They enrolled 200 high school students in a study where the control group (n=100) avoided all jellybeans for 2 months.
The treatment group (n=100) were given a free bag of jellybeans every week and were told to consume all of them before the next week.
At the end of the study, the acne level of each participant was scored.

```{r}
acne <- read.csv("../04-Data/acne.csv")
```

## Statistical Test

The scientists determined that there was no link between jelly beans and acne.

```{r}
t.test(Score ~ Group, data = acne)
ggplot(acne, aes(x=Group, y=Score)) + 
  geom_boxplot()
```

But people are skeptical and think that it might not be all jellybeans, but only those of a certain color.
So the scientists stop playing Minecraft long enough to conduct a second study. 
This time they have a control group (n=10) and twenty groups (n=10 per group) that each eat bags of only a single color of jellybean.

```{r}
acne2 <- read.csv("../04-Data/acne2.csv")

# a for loop probably would have been better!
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "purple"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "brown"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "pink"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "blue"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "teal"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "salmon"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "red"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "turquoise"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "magenta"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "yellow"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "grey"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "tan"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "cyan"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "green"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "mauve"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "beige"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "lilac"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "black"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "peach"))
t.test(Score ~ Color, data = filter(acne2, Color == "control" | Color == "orange"))
```

## Whoa

```{r}
ggplot(filter(acne2, Color == "control" | Color == "green"), aes(x=Color, y=Score)) + 
  geom_boxplot()
```

https://xkcd.com/882/

## Behind the numbers

Both acne datasets were generated using a random normal distribution.
```{r}
acne <- data.frame(round(rnorm(200,10,5)))
names(acne) = c("Score")
acne$Group <- c(rep("control", each=100), rep("jellybeans", each=100))

jellybeans = c("control","purple","brown","pink","blue","teal",
               "salmon","red","turquoise","magenta","yellow",
               "grey","tan","cyan","mauve","beige",
               "green","lilac","black","peach","orange")

acne2 <- data.frame(round(rnorm(210,10,5)))
names(acne2) = c("Score")
acne2$Color <- c(rep(jellybeans, each=10))
```

  
  