---
title: "Lab 03"
output: github_document
author: "Ken Field"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("NHANES")) install.packages("NHANES"); library(NHANES)
if (!require("UsingR")) install.packages("UsingR"); library(UsingR)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
```

## NHANES Dataset

Data from the US National Health and Nutrition Examination Study
https://www.cdc.gov/nchs/nhanes/index.htm

This is survey data collected by the US National Center for Health Statistics (NCHS) which has
conducted a series of health and nutrition surveys since the early 1960’s. Since 1999 approximately
5,000 individuals of all ages are interviewed in their homes every year and complete the health
examination component of the survey. The health examination is conducted in a mobile examination
center (MEC).

A description of the variables can be found in the package description.

```{r NHANES}
data("NHANES")
summary(NHANES)
help(NHANES)
```

## Model of Height Data

First look at the entire data set.

```{r Histogram}
ggplot(NHANES) +
  aes(x = Height) +
  geom_histogram(bins=100) +
  theme_cowplot()
```

Now lets just look at the data for children aged 2-17.

```{r Children Subset}
NHANES_child <- subset(NHANES, (Age >= 2 & Age <= 17)) #subset the full dataset
NHANES_child <- drop_na(NHANES_child, Height) #NA removal
ggplot(NHANES_child) +
  aes(x = Height) +
  geom_histogram(bins=100) +
  theme_cowplot()
```

## Model of the data
mean +- std deviation 
```{r Simple Model}
mean(NHANES_child$Height)
sd(NHANES_child$Height)
cat(paste('data', mean(NHANES_child$Height), '+/-', sd(NHANES_child$Height))) #poor model
summary(NHANES_child$Height) #This is the better model
```

The error is large when using mean to predict the height
Bad range of error with the first model (mean +/- std dev)
```{r}
error_mean <- NHANES_child$Height - mean(NHANES_child$Height)
ggplot(NULL, aes(error_mean)) +
  geom_histogram(bins = 100) + 
  xlim(-60, 60) +
  labs(
    x = "Error when predicting height with mean"
  )
rmse_mean <- sqrt(mean(error_mean**2))
rmse_mean 
```

More sophisticated models are necessary

```{r Age}
p1 <- NHANES_child %>% 
  ggplot(aes(x = Age, y = Height)) +
  geom_point(position = "jitter",size=0.05) +
  scale_x_continuous(breaks = seq.int(0, 20, 2)) +
  ggtitle('A: original data') +
  theme_cowplot()

lmResultHeightOnly <- lm(Height ~ Age + 0, data=NHANES_child)
lmResultHeightOnly
summary(lmResultHeightOnly)
rmse_heightOnly <- sqrt(mean(lmResultHeightOnly$residuals**2))
rmse_heightOnly

p2 <- NHANES_child %>% #whats %>%
  ggplot(aes(x = Age, y = Height)) +
  geom_point(position = "jitter",size=0.05) +
  scale_x_continuous(breaks = seq.int(0, 20, 2)) + 
  annotate('segment',x=0,xend=max(NHANES_child$Age),
           y=0,yend=max(lmResultHeightOnly$fitted.values),
           color='blue',lwd=1) + 
  ggtitle('B: age') +
  theme_cowplot()

lmResultHeight <- lm(Height ~ Age, data=NHANES_child) #implied constant in linear model (lm(Height ~ Age,...))
lmResultHeight
summary(lmResultHeight)
rmse_height <- sqrt(mean(lmResultHeight$residuals**2))
cat(paste('Root mean sq erro: ', rmse_height))

p3 <- NHANES_child %>% 
  ggplot(aes(x = Age, y = Height)) +
  geom_point(position = "jitter",size=0.05) +
  scale_x_continuous(breaks = seq.int(0, 20, 2)) + 
  geom_smooth(method='lm',se=FALSE) + 
  ggtitle('C: age + constant') +
  theme_cowplot()

plot_grid(p1,p2,p3,ncol=2) #plot several figures as one
```
This model is much better than using only the mean.
Dotplot by default adds jitter and bins to make data easier to see
As children get older they get taller
B: linear relationship (age/height), problem lowest age is 2 not 0, need to add a constant (not hard-coded)
Good model
Want RTMSE that is low
----------------------------------------------------

Can we make it even better?
Yes by looking at other factors contributing to height (create seperate fits for male and female)

```{r}
NHANES_child %>% 
  ggplot(aes(x = Age, y = Height)) +
  geom_point(aes(colour = factor(Gender)),  #differentiate male female
             position = "jitter", 
             alpha = 0.8,
             size=0.05) +
  geom_smooth(method='lm',aes(group = factor(Gender), 
                              colour = factor(Gender))) + 
  theme(legend.position = c(0.25,0.8)) + 
  ggtitle('age + constant + gender') +
  theme_cowplot()

lmResultGenderAge <- lm(Height ~ Gender + Age, data=NHANES_child)
lmResultGenderAge
summary(lmResultGenderAge) #r^2 closer to one is bueno
```

## Test Assumptions


Acknowledgements
=========================

```{r EZ print}
#to call EZ_print store the things to be printed in a variable (l = c(a, b, c, d...)) then call EZ_print(l)
EZ_print <- function(pirate) {
  return(cat(paste(pirate)))
  }
```



```{r}
a <- 42
b <- "hello"
c = c(a, b)
EZ_print(c)
```