---
title: "Lab 03"
output: github_document
author: "Ken Field"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if (!require("NHANES")) install.packages("NHANES"); library(NHANES)
if (!require("UsingR")) install.packages("UsingR"); library(UsingR)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
```

## NHANES Dataset

Data from the US National Health and Nutrition Examination Study
https://www.cdc.gov/nchs/nhanes/index.htm

This is survey data collected by the US National Center for Health Statistics (NCHS) which has
conducted a series of health and nutrition surveys since the early 1960’s. Since 1999 approximately
5,000 individuals of all ages are interviewed in their homes every year and complete the health
examination component of the survey. The health examination is conducted in a mobile examination
center (MEC).

A description of the variables can be found in the package description.

```{r NHANES}
data("NHANES")
summary(NHANES)
help(NHANES)
```

## Model of Height Data

First look at the entire data set.

```{r Histogram}
ggplot(NHANES) +
  aes(x = Height) +
  geom_histogram(bins=100) +
  theme_cowplot()
```

Now lets just look at the data for children aged 2-17.

```{r Children Subset}
NHANES_child <- subset(NHANES, (Age >= 2 & Age <= 17))
NHANES_child <- drop_na(NHANES_child, Height)
ggplot(NHANES_child) +
  aes(x = Height) +
  geom_histogram(bins=100) +
  theme_cowplot()
```

## Model of the data

```{r Simple Model}
mean(NHANES_child$Height)
sd(NHANES_child$Height)
summary(NHANES_child$Height)
```

The error is large when using mean to predict the height


```{r}
error_mean <- NHANES_child$Height - mean(NHANES_child$Height)
ggplot(NULL, aes(error_mean)) +
  geom_histogram(bins = 100) + 
  xlim(-60, 60) +
  labs(
    x = "Error when predicting height with mean"
  )
rmse_mean <- sqrt(mean(error_mean**2))
rmse_mean
```

More sophisticated models are necessary

```{r Age}
p1 <- NHANES_child %>% 
  ggplot(aes(x = Age, y = Height)) +
  geom_point(position = "jitter",size=0.05) +
  scale_x_continuous(breaks = seq.int(0, 20, 2)) +
  ggtitle('A: original data') +
  theme_cowplot()

lmResultHeightOnly <- lm(Height ~ Age + 0, data=NHANES_child)
lmResultHeightOnly
summary(lmResultHeightOnly)
rmse_heightOnly <- sqrt(mean(lmResultHeightOnly$residuals**2))
rmse_heightOnly

p2 <- NHANES_child %>% 
  ggplot(aes(x = Age, y = Height)) +
  geom_point(position = "jitter",size=0.05) +
  scale_x_continuous(breaks = seq.int(0, 20, 2)) + 
  annotate('segment',x=0,xend=max(NHANES_child$Age),
           y=0,yend=max(lmResultHeightOnly$fitted.values),
           color='blue',lwd=1) + 
  ggtitle('B: age') +
  theme_cowplot()

lmResultHeight <- lm(Height ~ Age, data=NHANES_child)
lmResultHeight
summary(lmResultHeight)
rmse_height <- sqrt(mean(lmResultHeight$residuals**2))
rmse_height

p3 <- NHANES_child %>% 
  ggplot(aes(x = Age, y = Height)) +
  geom_point(position = "jitter",size=0.05) +
  scale_x_continuous(breaks = seq.int(0, 20, 2)) + 
  geom_smooth(method='lm',se=FALSE) + 
  ggtitle('C: age + constant') +
  theme_cowplot()

plot_grid(p1,p2,p3,ncol=2)
```

This model is much better than using only the mean.

Can we make it even better?

```{r}
NHANES_child %>% 
  ggplot(aes(x = Age, y = Height)) +
  geom_point(aes(colour = factor(Gender)), 
             position = "jitter", 
             alpha = 0.8,
             size=0.05) +
  geom_smooth(method='lm',aes(group = factor(Gender), 
                              colour = factor(Gender))) + 
  theme(legend.position = c(0.25,0.8)) + 
  ggtitle('age + constant + gender') +
  theme_cowplot()

lmResultGenderAge <- lm(Height ~ Gender + Age, data=NHANES_child)
lmResultGenderAge
summary(lmResultGenderAge)
```

## Test Assumptions

```{r Data Exploration}

```

```{r t Test}
t.test(hatch.mass ~ sex, data = quail)
```

This shows us that hatch mass does depend on the sex of the chick.

After exploring the dataset, formulate some additional tentative hypotheses that you wish to test.
I have provided some examples of the types of graphs that you could consider. 
You should alter these examples for your own use (and fix any issues that find.)

```{r Data Visualization}

ggplot(quail) +
  aes(x = hatch.mass) +
  geom_histogram(bins=100) +
  theme_cowplot()

ggplot(quail) +
  aes(x = hatch.mass, color = sex, fill = sex) +
  geom_histogram(binwidth = 0.5, position = "dodge") +
  theme_cowplot()

ggplot(quail) +
  aes(x = hatch.mass,  fill = sex) + 
  geom_density(alpha=.3) +
  theme_cowplot()
```

There appears to be some other factors influencing the hatch mass, in addition to sex. 
Explore the data and determine how to test a hypothesis about one of these other factors.

```{r Statistical Tests}

```

Acknowledgements
=========================

