---
title: "Group Project 1"
subtitle: "Biology 368/664 Bucknell University"
output: html_notebook
author: Prof. Ken Field
date: 27 Jan 2020
---

This project will require you to develop a tutorial to teach students in one of the Biology, Neuroscience, or Animal Behavior courses how to use R for graphing and data analysis. 

## Classes

You will need to identify 1 or 2 other students to work with on the same tutorial. 
The course must be a 300-level course and you should plan on working with the instructor of that course to develop the tutorial.

```{r echo=FALSE}
biologyClasses <- read.csv("2019-2020BiologyCourses.csv", stringsAsFactors=FALSE)
print(biologyClasses)
```

## Groups

You will need to identify 1 or 2 other students to work with on the same tutorial. 
The course must be a 300-level course and you should plan on working with the instructor of that course to develop the tutorial.

```{r echo=FALSE}
Biol364_664roster <- read.csv("../00-Syllabus/Biol364_664roster.csv", stringsAsFactors=FALSE)
# Biol364_664roster$Course <- c("Biology 205",
#                               "Biology 206",
#                               "Biology 207",
#                               "Biology 206",
#                               "Biology 206",
#                               "A. Behavior 266",
#                               "Neuro 253",
#                               "Biology 208",
#                               "Biology 205",
#                               "Biology 208",
#                               "Biology 208",
#                               "Biology 207")
# summary(as.factor(Biol364_664roster$Course))
print(Biol364_664roster)
```

## Grading

Each student will be expected to complete the following tasks to earn 85% of the points available for this assignment (21/25).

- Identify and obtain suitable dataset
- Create a Github repository and use version control to collaborate on the project
- Spend 4-6 hours preparing, coding, and testing tutorial
  + Data exploration
  + Data visualization
  + Hypothesis testing
- Present tutorial in class
- Provide public archive suitable for sharing to students/faculty

Each group should use an *Acknowledgements* section to document the participation of each member and the collaboration within and between groups.

Additional credit will be awarded for providing assistance to other groups and for the development of a tutorial that goes beyond the minimal expectations listed above.


