---
title: "Han homework 1"
author: "Robert Han"
date: "1/13/2020"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install necessary libraries (and maybe some unnecessary ones I'm not sure.)
```{r}
if (!require("UsingR")) install.packages("UsingR"); library(UsingR)
if (!require("cowplot")) install.packages("cowplot"); library(cowplot)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse)
if (!require("readxl")) install.packages("readxl"); library(readxl)
if (!require("gplots")) install.packages("gplots"); library(gplots) 
```


Import and read data using readxl. Provide summary statistics for the data. This code block also replaces the
binary representations of hibernation and diet data with character ones.
https://stackoverflow.com/questions/5824173/replace-a-value-in-a-data-frame-based-on-a-conditional-if-statement
```{r data summary}
project_data <- read_excel("Supplementary material.xlsx")

library(plyr)
project_data$hibernation <- as.character(project_data$hibernation)
project_data$hibernation <- revalue(project_data$hibernation, c("0" = "Hibernator", "1" = "Nonhibernator"))

project_data$diet <- as.character(project_data$diet)
project_data$diet <- revalue(project_data$diet, c("1" = "herbivore or folivore", "2" = "frugivore/folivore or granivore", "3" = "frugivore/faunivore or omnivore", "4" = "faunivore, piscivore, carnivore or insectivore"))

summary(project_data)
```

From these summary statistics we can get a general sense for what data is present in the dataset. For example, we
can tell that there are 1104 entries in the dataset as well as which variables are likely continuous or
categorical in nature.

The dataset has the given variables:

Group: Cat
Genus species: Cat 
brain mass: Cont
Ref. brain mass: Null/Cat
body mass: Cont
Ref. body mass: Null/Cat
hibernation: Cat
Ref. hibernation: Null/Cat
diurnality: Cat
Ref. diurnality: Null/Cat
substrate use: Cat
Ref. substrate use: Null/Cat
diet: Cat
Ref diet: Null/Cat
mid-latitude of geographical distribution: ?
Ref. mid-latitude of geographical distribution: Null/Cat

To take a closer look at the data entries we can look at the data. As a result of the earlier data manipulations
some of the variables have been changed from binary representations to character representations.
```{r Visualize data as table}
project_data
```

To get a sense for the distribution of data entries across their encompassing groups we can visualize counts of
the later in a bar plot. Due to the number of groups represented in the dataset (n = 25) I found it was better 
to visualize the bar plot horizontally. From this visualization its clear not all groups are equally represented
in the dataset. Rodentia, Primates, and Carnivora in particular make up close to 70% of all the species sampled. 

https://www.statmethods.net/graphs/bar.html
https://www.datamentor.io/r-programming/bar-plot/
```{r Barplot of groups}
ggplot(project_data) +
  aes(x = Group) +
  geom_bar() +
  ylab("Count") +
  coord_flip() +
  theme_cowplot()
```

At a glance the Q-Q plots suggest the data is not normally distributed. Using log10 transformations appeared to
enhance the Q-Q plot slightly. The idea of creating variables to hold column data came from Fallon.
```{r qq test}
brain = project_data$`brain mass`
body = project_data$`body mass`

qqnorm(brain, main='Normal Q-Q Plot for brain mass')
qqline(brain)

qqnorm(log(brain), main='Normal Q-Q Plot for log10 brain mass')
qqline(log(brain))

qqnorm(body, main='Normal Q-Q Plot for body mass')
qqline(body)

qqnorm(log(body), main='Normal Q-Q Plot for log10 body mass')
qqline(log(body))

qqnorm(brain/body, main='Normal Q-Q Plot for brain/body mass')
qqline(brain/body)

qqnorm(log(brain/body), main='Normal Q-Q Plot for log10(brain/body mass)')
qqline(log(brain/body))
```

To look more carefully at each of the quantitative variables, a simple eda was used for brain mass and brain 
mass relative to body mass. A log10 transformation of brain mass showed less skew and better Q-Q plot. Log10
transformation of relative brain mass did not reduce skew very much; however resulted in a better Q-Q plot.
```{r simple exploratory data analysis}
simple.eda(project_data$`brain mass`)
simple.eda(log(project_data$`brain mass`))
simple.eda(project_data$`body mass`)
simple.eda(log(project_data$`body mass`))
simple.eda(project_data$`brain mass`/project_data$`body mass`)
simple.eda(log(project_data$`brain mass`/project_data$`body mass`))
```

From the Shapiro-Wilk normality test we get p-values much smaller than 0.05. According to Fallon (she remembers
more stats than I do) we would want the p value to be larger than 0.05; since it is not, I believe we assume the
data is not normally distributed. This conclusion aligns with with the exploratory data analysis done beforehand
where the data did not appear very normal even after log10 transformations.
```{r}
shapiro.test(project_data$`brain mass`)
shapiro.test(project_data$`body mass`)
shapiro.test(project_data$`brain mass`/project_data$`body mass`)
```

Show counts of hibernators and nonhibernators as a table. We can see there are many
hibernators in the dataset. Then again we should probably visualize it in a graphic.
```{r Visualize subset of data}
table(project_data$hibernation)
```

Visualize counts of hibernators and nonhibernators as a barplot. There are many hibernators! Though it is
somewhat difficult to differentiate some of the coloring, we can still see that a few groups are overrepresented
in the dataset. 
```{r Hibernation barplot}
ggplot(project_data) + 
  geom_bar(aes(x = `hibernation`, fill = Group)) + 
  xlab("Hibernation") + 
  ylab("Count") +
  theme_cowplot()
```

Scatterplot of brain mass and body mass after log10 transformation. The log10 transformation suggests a trend
between brain mass and body mass. The hypothesis being tested is whether or not hibernation and brain size are
related. This visualization is not directly tied to the hypothesis but dang is it nice to look at. Interestingly,
after the log transformation we see a greater concentration of nonhibernators at the tail of the graph. The
linear model is very small and not immediately seen. No units are provided in the dataset.
```{r}
ggplot(project_data) +
  aes(y = log(`brain mass`), x = log(`body mass`)) + 
  theme(legend.position = c(0.05, 0.8)) +
  geom_point(aes(color = hibernation), alpha = 0.5) +
  geom_smooth(method=lm , color="black", se=TRUE) +
  ylab("log10 Brain mass") + 
  xlab("log10 Body mass") +
  theme_cowplot()
```

The results of the linear regression statistics. The p-value calculated is well under 0.05 which is not really
relevant to the hypothesis. 
```{r Statistics on linear regression}
mass_fit = lm(log(`brain mass`)~log(`body mass`), data=project_data)
summary(mass_fit)
```

The boxplot was generated to better visualize the differences between the brain mass of hibernators and
nonhibernators. a log10 transformation was used again to better see the data. The boxplot suggests the median
brain size of hibernating species is actually larger than that of nonhibernators. 
```{r Brain mass viewed as a boxplot}
ggplot(project_data) +
  aes(x = `hibernation`, y = log(`brain mass`)) +
  geom_boxplot() +
  guides(alpha=FALSE) +
  xlab("Hibernation") + 
  ylab("log10 Brain mass") +
  theme_cowplot()
```

It may be helpful to see the actual data points so a jitter overlay was added.
```{r Brain mass viewed as a boxplot with jitter overlay}
ggplot(project_data) +
  aes(x = `hibernation`, y = log(`brain mass`)) +
  geom_boxplot() +
  geom_point(alpha = 0.3, position = "jitter", colour = "blue") +
  guides(alpha=FALSE) +
  xlab("Hibernation") + 
  ylab("log10 Brain mass") +
  theme_cowplot()
```

Accounting for body mass the boxplot suggests the difference in brain size between hibernators and nonhibernators
may not be as large as the previous figure suggests. A different function for generating the boxplot was used on
a whim. A log10 transformation was used since it made the eda data look nicer. Unsurprisingly the relative brain
mass was a low value resulting in negative values after log transformation.

https://www.statmethods.net/graphs/boxplot.html
```{r brain mass/body mass ~ hibernation boxplot}
boxplot(log(`brain mass` / `body mass`) ~ hibernation , data = project_data, 
        xlab = "Hibernation", ylab = "Relative brain mass")
```

Looking directly at the mean brain mass/relative brain mass suggests nonhibernators have a smaller brain mass
than hibernators; however, when accounting for body size nonhibernators may have a greater relative brain mass
than hibernators.

http://www.sthda.com/english/wiki/plot-group-means-and-confidence-intervals-r-base-graphs
```{r}
plotmeans(log(`brain mass`) ~ `hibernation`, data = project_data, xlab = "Diet", ylab = "Log10 brain mass")

plotmeans(`brain mass`/`body mass` ~ `hibernation`, data = project_data, xlab = "Diet", ylab = "Relative brain mass")
```

Perform T-test to determine if the difference in means is significant. According to the T-test results, the
difference in means is significant. With regard to the primary hypothesis being tested, hibernators have a
significantly lower brain mass relative to body mass when compared to nonhibernators; however hibernators have a
significantly greater brain mass in general when compared to nonhibernators.

https://www.statmethods.net/stats/ttest.html
```{r T-test}
t.test(log(`brain mass`) ~ `hibernation`, data = project_data)

t.test((`brain mass`/`body mass`) ~ `hibernation`, data = project_data)
```

Alternative hypothesis: Predatory (carnivore, insectivore, piscivore, faunivore) diets enhances brain size
evolution.

The barplot shows a relatively even distribtion of diets amongst the species sampled. 
For comparison to the original hypothesis we also see that frugivores/foliovores or granivores have a greater
ratio of nonhibernators to hibernators than any other diet group. If the former hypothesis were true we might
expect the carnivore/piscivore/insectivore diet to have the greatest brain mass given the greater number of
nonhibernators in that category.
```{r barplot diet}
ggplot(project_data) +
  aes(x = diet, fill = hibernation) +
  geom_bar() +
  theme_cowplot()
```

Species which feed on other animals showed a higher maximum relative brain mass than other diet types. That being
said, the median and interquartile range of said diet was nothing specatacular and even fell behind two plant
based diets. I would have liked to show the means of each diet on the boxplot but was unable to figure that one
out. I also chose to visualize a boxplot of the log transformed brain mass. Interestingly the boxplots suggests
preying on other animals is associated with less brain mass than a strictly plant based diet. The herbivore or
folivore diet in particular appeared to be linked to animals with the greatest brain mass.
```{r boxplots of brain mass and relative brain mass}
boxplot(log(`brain mass`) ~ diet , data = project_data,
        xlab = "Diet", ylab = " log10 Brain mass")

boxplot(`brain mass` / `body mass` ~ diet , data = project_data, 
        xlab = "Diet", ylab = "Relative brain mass (units)")
```

Since the range and quartiles of the data was visualized earlier, I thought a good next step would be to see the
means, especially given that I couldnt figure out how to overlay them onto the boxplot. We can see
herbivores/foliovores have the greatest average brain mass but the lowest relative brain mass. The
carnivore/piscivore/insectivore diet actually had the smallest average brain mass which contradicts my
hypothesis. These observations align with what was shown in the boxplots.

```{r Mean brain mass and relative brain mass}
# Plot the mean of teeth length by dose groups
plotmeans(log(`brain mass`) ~ `diet`, data = project_data, xlab = "Diet", ylab = " log10 Brain mass")

plotmeans((`brain mass` / `body mass`) ~ `diet`, data = project_data, xlab = "Diet", ylab = " Relative brain mass")
```

Not sure if this is the appropriate test. Obviously the datset used was not normally distributed as discussed
earlier. The calculated p-values are much smaller than 0.05 suggesting the differences between some of the means
are statistically significant. The difference between herbivore/folivore and any other diet is certainly
noticeable. Regardless, the alternative hypothesis is definitely not supported by the dataset on hand. Predatory
(carnivore, insectivore, piscivore, faunivore) diets do not appear to be associated with increased brain size
whether it be in straight brain mass or relative brain mass.

https://www.r-bloggers.com/performing-anova-test-in-r-results-and-interpretation/
```{r}
res1.aov <- aov(`brain mass` ~ diet, data = project_data)
summary(res1.aov)

res2.aov <- aov(`brain mass`/`body mass` ~ diet, data = project_data)
summary(res2.aov)
```




"A DEFINITE TREND IS EVIDENT"
https://twitter.com/mantyne/status/1217444724656701440