---
title: "Homework 7"
author: "Biology Student"
date: "1 Apr 2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries

```{r Load Libraries, include=FALSE}
if (!require("devtools")) install.packages("devtools"); library(devtools)
if (!require("tidyverse")) install.packages("tidyverse"); library(tidyverse) # Load tidyverse last
```

## Prerequisites for Homework 7

1. Completion of the [Galaxy](1-Galaxy.md) and [SARTools](2-SARTools.md) tutorials.
2. Results from either EdgeR or DESeq2 analysis of differential **gene** expression of the Pasilla dataset (after batch effect correction)
3. Results from the EdgeR and DESeq2 analysis of differential **transcript** expression of the Pasilla dataset (after batch effect correction)

## Batch Correction in SARTools

As we noted during the previous lab, the Single and Paired samples were quite different in the Pasilla dataset.
This difference was a problem because it was not related to the hypothesis: that transcript expression would be affected by RNAi treatment.
For this reason, we want to block the effects of sequencing layout on the experimental design by using batch correction.
In SARTools this is done simply by providing a column in the target.txt file that contains the batch information and then setting `batch` to that column name.
You should run the SARTools analysis again after setting
batch <- "batch"
and changing the working directory to a new directory (like "SARTools.DESeq2.genes.batch"). 
You will need to also make that directory in the Files panel.
Lastly, you should probably change the project name to something like "SARTools.DESeq2.genes.batch".

## Transcript Expression

Because we expect Pasilla knockdown to affect alternative splicing, we need to look at transcript expression and not just gene expression.
Run the DESeq2 and edgeR analyses using transcripts instead of genes. (You will still need to use batch correction.)
Don't forget to make new directories for the output and give the project a new name so that the previous files don't get overwritten.

## Objectives

[ ] Compare the number of genes differentially expressed before and after batch correction
[ ] Confirm whether Pasilla gene expression was knocked down by the RNAi treatment
[ ] Compare the differential transcript expression detected by edgeR and DESeq2

## Batch Correction

Compare the SARTools Report for one of the analyses with and without batch correction.
In one or two sentences, summarize the difference that batch correction made in the PCA plot and the number of differentially expressed genes.

## Pasilla Gene Analysis

1. Use http://flybase.org/ to determine the FlyBase **gene** name for the Pasilla gene.

You are looking for the "Flybase ID".

2. Load in the gene expression results from either EdgeR or DESeq2.

Below I demonstrate how to load in both, but you only need to use one or the other.

```{r Loading in Data}
# DESeq2
DESeq2_genes_complete <- read_delim("SARTools/SARTools.DESeq2.genes/tables/RNAivsUntreated.complete.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
# Check that numbers agree with SARTools Report
DESeq2_genes_complete %>% dplyr::filter(padj < 0.05) %>% tally()
DESeq2_genes_complete %>% dplyr::filter(padj < 0.05) %>% dplyr::filter(FoldChange > 1) %>% tally()
DESeq2_genes_complete %>% dplyr::filter(padj < 0.05) %>% dplyr::filter(FoldChange < 1) %>% tally()

# edgeR
edgeR_genes_complete <- read_delim("SARTools/SARTools.edgeR.genes/tables/RNAivsUntreated.complete.txt", 
    "\t", escape_double = FALSE, trim_ws = TRUE)
# Check that numbers agree with SARTools Report
edgeR_genes_complete %>% dplyr::filter(padj < 0.05) %>% tally()
edgeR_genes_complete %>% dplyr::filter(padj < 0.05) %>% dplyr::filter(FC > 1) %>% tally()
edgeR_genes_complete %>% dplyr::filter(padj < 0.05) %>% dplyr::filter(FC < 1) %>% tally()
```

3. Determine if the Pasilla gene was differentially expressed in the treated samples.

```{r Loading in Data}
# Looking for a specific gene
DESeq2_genes_complete %>% dplyr::filter(Id == "FBgn0261552")
edgeR_genes_complete %>% dplyr::filter(Id == "FBgn0261552")
```

What is the expression level (normalized counts) for each treatement group?
Untreated:
RNAi-treated:

What is the fold-change difference in expression level between the two treatment groups?


How significant is this difference in expression level?


## Comparison of EdgeR and DESeq2

1. Load in the transcript expression results from both EdgeR and DESeq2.

```{r Load transcript results}

```

Note that DESeq2 and edgeR have the same raw counts (as expected), but different counts after normalization and different Differential Expression results.

2. Determine the number of differentially expressed transcripts detected by each program.

Separately tally the significantly (FDR < 0.05) upregulated and downregulated transcripts for each analysis pipeline.

```{r}

```

3. Compare the identity of the transcripts differentially expressed

Do both programs identify exactly the same genes as differentially expressed?

I suggest using the `intersect()` and `setdiff()` functions to compare the list of Ids for each subset of transcripts.

https://stat.ethz.ch/R-manual/R-devel/library/base/html/sets.html

```{r}

```


Bonus, if you are feeling brave. Look for a new package that allows you to make Venn diagrams of the differentially expressed transcripts. 

## Helpful RNA-Seq Links

Pasilla paper: https://genome.cshlp.org/content/21/2/193.full

RNA-seqlopedia: https://rnaseq.uoregon.edu/

RNA-Seq Blog: http://www.rna-seqblog.com/

QCFAIL Blog: https://sequencing.qcfail.com/
(Unfortunately it looks like they are no longer posting, but they have some great posts about problems with certain Illumina sequencers.)

QCFAIL post about SRA file corruption: https://sequencing.qcfail.com/articles/data-can-be-corrupted-upon-extraction-from-sra-files/ 
(This is why it is so important to look at the raw fastq files and check the lengths of the reads before trimming.)
